---
title: "Does analytics work in professional sports?"
author: "Walt DeGrange"
date: "November 11, 2020"
output:
  html_document: default
  pdf_document: default
---

==

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE ,warning = FALSE,message = FALSE)
```

```{r Libraries}
library(dplyr)
library(tidyr)
library(magrittr)
library(ggplot2)
library(purrr)
library(ggforce)
library(fpc)
library(cluster) 
library(mclust)
library(formattable)
library(formatR)
library(officer)
library(flextable)
# library(lares)
```

```{r Import Presentation}

UofArkPres <- read_pptx("Base File - U of Ark.pptx")

```

# Sports Analytics - Does it make a difference?

Salary Data - <https://www.spotrac.com>

```{r Data Import}
TeamData <- read.csv("ESPNSportsAnalytics.csv")

colnames(TeamData)[1] <- "Sport"

```

```{r Create data tables, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

TeamData %<>% mutate(PopulationMil = digits(Market/1000000,1))

TeamDataKM2015 <- TeamData %>% select(X2014HF,X2015HF)

TeamDataKM2016 <- TeamData %>% select(X2014HF,X2015HF,X2016HF)

TeamDataKM2017 <- TeamData %>% select(X2014HF,X2015HF,X2016HF,X2017HF)

TeamDataKM2018 <- TeamData %>% select(X2014HF,X2015HF,X2016HF,X2017HF,X2018HF)

TeamDataKM2019 <- TeamData %>% select(X2014HF,X2015HF,X2016HF,X2017HF,X2018HF,X2019HF)

TeamDataKM2020 <- TeamData %>% select(X2014HF,X2015HF,X2016HF,X2017HF,X2018HF,X2019HF,X2020HF)

```

## Overall

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

TeamDataNew <- TeamData %>% 
  group_by(Team) %>% 
  mutate(HowFar = digits(sum(X2014HF,X2015HF,X2016HF,X2017HF,X2018HF,X2019HF,X2020HF)/7,2))

# New team only last three years with 8 total for how far
TeamDataNew$HowFar[TeamDataNew$Team == "Golden Knights"] <- 8/3

TeamDataNewDT <- TeamDataNew %>% 
  group_by(Sport) %>% 
  summarise(count = n(), MeanESPN = digits(mean(ESPNRating),1), StdESPN = digits(sd(ESPNRating),1)) %>% 
  flextable() %>%
  bold(part = "header") %>% 
  align(align = "center", part = "header" ) %>% 
  fontsize(size = 16, part = "all") %>% 
  autofit() %>% 
  bg(bg = "white",part = "all")

UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "Overview By Sport", location = ph_location_type(type = "title")) %>% 
  ph_with(TeamDataNewDT, location = ph_location(left = 3, top = 2))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 1)

TeamDataNewDT

```

## Local Teams

```{r Local Teams, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

LocalTeams <- TeamDataNew %>% 
  filter(City %in% c("Dallas", "Kansas City", "Memphis")) %>% 
  group_by(City,Sport,Team,HowFar)  %>% 
  summarise(count = n(), ESPN = ESPNRating) %>% 
  select(City,Team,ESPN,HowFar) %>% 
  flextable() %>%
  colformat_num(j = 5, digits = 1) %>% 
  bold(part = "header") %>% 
  align(align = "center", part = "header" ) %>% 
  fontsize(size = 16, part = "all") %>% 
  autofit() %>% 
  bg(bg = "white",part = "all")

UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "Local Teams", location = ph_location_type(type = "title")) %>% 
  ph_with(LocalTeams, location = ph_location(left = 2.5, top = 2))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

LocalTeams
```

------------------------------------------------------------------------

# Champions

## 2014

```{r 2014, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

Champ2014 <- TeamDataNew %>% 
  filter(X2014HF == "5") %>%  
  select(Sport,Team,City,ESPNRating,PopulationMil) %>%  
  arrange(Sport) %>% 
  head(10) %>%  
  flextable() %>%
  bold(part = "header") %>% 
  align(align = "center", part = "header" ) %>% 
  fontsize(size = 16, part = "all") %>% 
  autofit() %>% 
  bg(bg = "white",part = "all")


UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "Champions 2014", location = ph_location_type(type = "title")) %>% 
  ph_with(Champ2014, location = ph_location(left = 1.5, top = 2))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

Champ2014

```

## 2015

```{r 2015, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

Champ2015 <- TeamDataNew %>% 
  filter(X2015HF == "5") %>%  
  select(Sport,Team,City,ESPNRating,PopulationMil) %>%  
  arrange(Sport) %>% 
  head(10) %>%  
  flextable() %>%
  bold(part = "header") %>% 
  align(align = "center", part = "header" ) %>% 
  fontsize(size = 16, part = "all") %>% 
  autofit() %>% 
  bg(bg = "white",part = "all")


UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "Champions 2015", location = ph_location_type(type = "title")) %>% 
  ph_with(Champ2015, location = ph_location(left = 1.5, top = 2))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

Champ2015

```

## 2016

```{r 2016, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

Champ2016 <- TeamDataNew %>% 
  filter(X2016HF == "5") %>%  
  select(Sport,Team,City,ESPNRating,PopulationMil) %>%  
  arrange(Sport) %>% 
  head(10) %>%  
  flextable() %>%
  bold(part = "header") %>% 
  align(align = "center", part = "header" ) %>% 
  fontsize(size = 16, part = "all") %>% 
  autofit() %>% 
  bg(bg = "white",part = "all")


UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "Champions 2016", location = ph_location_type(type = "title")) %>% 
  ph_with(Champ2016, location = ph_location(left = 1.5, top = 2))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

Champ2016

```

## 2017

```{r 2017, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

Champ2017 <- TeamDataNew %>% 
  filter(X2017HF == "5") %>%  
  select(Sport,Team,City,ESPNRating,PopulationMil) %>%  
  arrange(Sport) %>% 
  head(10) %>%  
  flextable() %>%
  bold(part = "header") %>% 
  align(align = "center", part = "header" ) %>% 
  fontsize(size = 16, part = "all") %>% 
  autofit() %>% 
  bg(bg = "white",part = "all")


UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "Champions 2017", location = ph_location_type(type = "title")) %>% 
  ph_with(Champ2017, location = ph_location(left = 1.5, top = 2))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

Champ2017

```

## 2018

```{r 2018, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

Champ2018 <- TeamDataNew %>% 
  filter(X2018HF == "5") %>%  
  select(Sport,Team,City,ESPNRating,PopulationMil) %>%  
  arrange(Sport) %>% 
  head(10) %>%  
  flextable() %>%
  bold(part = "header") %>% 
  align(align = "center", part = "header" ) %>% 
  fontsize(size = 16, part = "all") %>% 
  autofit() %>% 
  bg(bg = "white",part = "all")


UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "Champions 2018", location = ph_location_type(type = "title")) %>% 
  ph_with(Champ2018, location = ph_location(left = 1.5, top = 2))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

Champ2018

```

## 2019

```{r 2019, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

Champ2019 <- TeamDataNew %>% 
  filter(X2019HF == "5") %>%  
  select(Sport,Team,City,ESPNRating,PopulationMil) %>%  
  arrange(Sport) %>% 
  head(10) %>%  
  flextable() %>%
  bold(part = "header") %>% 
  align(align = "center", part = "header" ) %>% 
  fontsize(size = 16, part = "all") %>% 
  autofit() %>% 
  bg(bg = "white",part = "all")


UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "Champions 2019", location = ph_location_type(type = "title")) %>% 
  ph_with(Champ2019, location = ph_location(left = 1.5, top = 2))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

Champ2019

```

## 2020

```{r 2020, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

Champ2020 <- TeamDataNew %>% 
  filter(X2020HF == "5") %>%  
  select(Sport,Team,City,ESPNRating,PopulationMil) %>%  
  arrange(Sport) %>% 
  head(10) %>%  
  flextable() %>%
  bold(part = "header") %>% 
  align(align = "center", part = "header" ) %>% 
  fontsize(size = 16, part = "all") %>% 
  autofit() %>% 
  bg(bg = "white",part = "all")


UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "Champions 2020", location = ph_location_type(type = "title")) %>% 
  ph_with(Champ2020, location = ph_location(left = 1.5, top = 2))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

Champ2020

```

------------------------------------------------------------------------

## Team list grouped by ESPN Ranking

```{r ESPN Rank, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

TeamDataNew$PopulationMil <- digits(TeamDataNew$PopulationMil,1)

ESPNRank <- TeamDataNew %>% 
  group_by(ESPNRating) %>% 
  summarise(Count = n(), HowFar = digits(sum(X2014HF,X2015HF,X2016HF,X2017HF,X2018HF,X2019HF,X2020HF)/(7*n()),1), MeanPop = digits(mean(PopulationMil),1), MeanSalary = digits(sum(X2014Salary,X2015Salary,X2016Salary,X2017Salary,X2018Salary,X2019Salary,X2020Salary)/(7*n()),1)) %>% 
  flextable() %>%
#  colformat_num(j = 5, digits = 1) %>% 
  bold(part = "header") %>% 
  align(align = "center", part = "header" ) %>% 
  fontsize(size = 16, part = "all") %>% 
  autofit() %>% 
  bg(bg = "white",part = "all")

UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "ESPN Rating", location = ph_location_type(type = "title")) %>% 
  ph_with(ESPNRank, location = ph_location(left = 2, top = 2))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

ESPNRank

```

## Ten best ESPN Rating performers

```{r ESPN Best Performers, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

ESPNBest <- TeamDataNew %>% 
  select(Sport,Team,City,ESPNRating,PopulationMil,HowFar) %>% 
  arrange(-HowFar) %>% 
  head(10, ESPNRating)  %>% 
  flextable() %>%
  colformat_num(j = 6, digits = 1) %>% 
  bold(part = "header") %>% 
  align(align = "center", part = "header" ) %>% 
  fontsize(size = 16, part = "all") %>% 
  autofit() %>% 
  bg(bg = "white",part = "all")

UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "ESPN Ranking Best Performance", location = ph_location_type(type = "title")) %>% 
  ph_with(ESPNBest, location = ph_location(left = 1, top = 2))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

ESPNBest

```

## Ten bottom ESPN Rating performers

```{r Bottom 10, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

ESPNWorst <- TeamDataNew %>% 
  select(Sport,Team,City,ESPNRating,PopulationMil,HowFar) %>%  
  arrange(-ESPNRating,HowFar) %>%
  head(10) %>%  
  flextable() %>%
  colformat_num(j = 6, digits = 1) %>% 
  bold(part = "header") %>% 
  align(align = "center", part = "header" ) %>% 
  fontsize(size = 16, part = "all") %>% 
  autofit() %>% 
  bg(bg = "white",part = "all")

UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "ESPN Ranking Worst Performance", location = ph_location_type(type = "title")) %>% 
  ph_with(ESPNWorst, location = ph_location(left = 1, top = 2))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

ESPNWorst
```

------------------------------------------------------------------------

## Big to Small Market Teams

### Los Angeles

```{r Big to Small Market, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

LosAngeles <- TeamDataNew %>% 
  group_by(Sport,Team,HowFar) %>% 
  filter(City %in% c("Los Angeles","Anaheim")) %>%  
  summarise(count = n(), ESPN = (ESPNRating)) %>% 
  select(Team,ESPN,HowFar) %>% 
  flextable() %>%
  colformat_num(j = 4, digits = 1) %>% 
  bold(part = "header") %>% 
  align(align = "center", part = "header" ) %>% 
  fontsize(size = 16, part = "all") %>% 
  autofit() %>% 
  bg(bg = "white",part = "all")

UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "Los Angeles Team Performance", location = ph_location_type(type = "title")) %>% 
  ph_with(LosAngeles, location = ph_location(left = 3, top = 2))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

LosAngeles

```

### New York

```{r New York, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

NewYork <- TeamDataNew %>% 
  group_by(Sport,Team,HowFar) %>% 
  filter(City %in% c("New York")) %>%  
  summarise(count = n(), ESPN = (ESPNRating)) %>% 
  select(Team,ESPN,HowFar) %>% 
  flextable() %>%
  colformat_num(j = 4, digits = 1) %>% 
  bold(part = "header") %>% 
  align(align = "center", part = "header" ) %>% 
  fontsize(size = 16, part = "all") %>% 
  autofit() %>% 
  bg(bg = "white",part = "all")

UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "New York Team Performance", location = ph_location_type(type = "title")) %>% 
  ph_with(NewYork, location = ph_location(left = 3, top = 2))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

NewYork

```

### Small Cities

```{r Green Bay, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

SmallCity <- TeamDataNew %>% 
  group_by(Sport,Team,HowFar,PopulationMil) %>% 
  filter(City %in% c("Green Bay", "Sunrise")) %>%  
  summarise(count = n(), ESPN = (ESPNRating)) %>% 
  select(Team,ESPN,HowFar,PopulationMil) %>% 
  flextable() %>%
  colformat_num(j = 4, digits = 1) %>% 
  bold(part = "header") %>% 
  align(align = "center", part = "header" ) %>% 
  fontsize(size = 16, part = "all") %>% 
  autofit() %>% 
  bg(bg = "white",part = "all")

UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "Small City's Team Performance", location = ph_location_type(type = "title")) %>% 
  ph_with(SmallCity, location = ph_location(left = 2, top = 2))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

SmallCity

```

## Plot of ESPN vs How Far a team made it into the playoffs on average for the 4 years of data

```{r Plots, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

TeamDataNew %>% ggplot(aes(x=ESPNRating,y=HowFar)) +
  geom_point()+
  geom_smooth()


```

## K Means clustering 2020

```{r KMeans Cluster 2019, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

TeamDataKM2020 <- scale(TeamDataKM2020)

fit <- kmeans(TeamDataKM2020, 2)

TeamDataKM2020 <- data.frame(TeamDataKM2020, fit$cluster)

TeamData2020 <- data.frame(TeamData, fit$cluster)

rownames(TeamData2020) <- paste(TeamData2020$Sport,TeamData2020$Team,sep="")


clusplot(TeamData2020, fit$cluster, color=TRUE, shade=FALSE, 
    labels=2, lines=0, cex.txt = .5,col.clus = c("red", "green"),col.txt = c("black", "black"))

TeamData2020 %<>% rename(Cluster = fit.cluster) 

TeamData2020DT <- TeamData2020 %>% 
  group_by(Cluster) %>% 
  summarise(Teams=n(),MeanX2014HF = digits(mean(X2014HF),1),MeanX2015HF = digits(mean(X2015HF),1),MeanX2016HF = digits(mean(X2016HF),1),MeanX2017HF = digits(mean(X2017HF),1),MeanX2018HF = digits(mean(X2018HF),1),MeanX2019HF = digits(mean(X2019HF),1),MeanX2020HF = digits(mean(X2020HF),1),MeanPopulation = digits(mean(PopulationMil),1),MeanSalary = digits(mean(X2020Salary),1),MeanESPNRating = digits(mean(ESPNRating),1))

# TeamData2019DT <- as.data.frame(t(TeamData2019DT))
# 
# colnames(TeamData2019DT) <- c("1","2")

TeamData2020DTFT <- TeamData2020DT %>% 
  flextable() %>%
#  colformat_num(digits = 1) %>% 
  bold(part = "header") %>% 
#  align(align = "center", part = "header" ) %>% 
  fontsize(size = 12, part = "all") %>% 
  rotate(rotation = "tbrl", align = "center", part = "header") %>% 
  align(align = "right", part = "header") %>% 
  valign(valign = "center", part = "header") %>% 
  align(align = "center", part = "body") %>% 
  set_table_properties(width = .2, layout = "autofit") %>% 
#  autofit() %>% 
  height(height = 2, part = "header") %>% 
  bg(bg = "white",part = "all") %>% 
  bg(j = 12, bg = "yellow", part = "all") %>% 
  bg(j = 10:11, bg = "green", part = "all")

UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>%
  ph_with(value = "2020 K-Means Clustering Plot", location = ph_location_type(type = "title")) %>%
  ph_with(external_img("km2020plot.png"), location = ph_location_type(type = "body"), use_loc_size = TRUE)

UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
  ph_with(value = "2020 K-Means Clustering Data", location = ph_location_type(type = "title")) %>% 
  ph_with(TeamData2020DTFT, location = ph_location(left = 0.8, top = 2.5))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

TeamData2020DTFT

```

## K Means clustering 2019

```{r KMeans Cluster 2019, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

TeamDataKM2019 <- scale(TeamDataKM2019)

fit <- kmeans(TeamDataKM2019, 2)

TeamDataKM2019 <- data.frame(TeamDataKM2019, fit$cluster)

TeamData2019 <- data.frame(TeamData, fit$cluster)

rownames(TeamData2019) <- paste(TeamData2019$Sport,TeamData2019$Team,sep="")

clusplot(TeamData2019, fit$cluster, color=TRUE, shade=FALSE, 
  	labels=2, lines=0, cex.txt = .5,col.clus = c("red", "green"),col.txt = c("black", "black"))

TeamData2019 %<>% rename(Cluster = fit.cluster) 

TeamData2019DT <- TeamData2019 %>% 
  group_by(Cluster) %>% 
  summarise(Teams=n(),MeanX2014HF = digits(mean(X2014HF),1),MeanX2015HF = digits(mean(X2015HF),1),MeanX2016HF = digits(mean(X2016HF),1),MeanX2017HF = digits(mean(X2017HF),1),MeanX2018HF = digits(mean(X2018HF),1),MeanX2019HF = digits(mean(X2019HF),1),MeanPopulation = digits(mean(PopulationMil),1),MeanSalary = digits(mean(X2019Salary),1),MeanESPNRating = digits(mean(ESPNRating),1))

# TeamData2019DT <- as.data.frame(t(TeamData2019DT))
# 
# colnames(TeamData2019DT) <- c("1","2")

TeamData2019DTFT <- TeamData2019DT %>% 
  flextable() %>%
#  colformat_num(digits = 1) %>% 
  bold(part = "header") %>% 
#  align(align = "center", part = "header" ) %>% 
  fontsize(size = 12, part = "all") %>% 
  rotate(rotation = "tbrl", align = "center", part = "header") %>% 
  align(align = "right", part = "header") %>% 
  valign(valign = "center", part = "header") %>% 
  align(align = "center", part = "body") %>% 
  set_table_properties(width = .2, layout = "autofit") %>% 
#  autofit() %>% 
  height(height = 1.5, part = "header") %>% 
  bg(bg = "white",part = "all") %>% 
  bg(j = 11, bg = "yellow", part = "all") %>% 
  bg(j = 9:10, bg = "green", part = "all")

# UofArkPresNew <- UofArkPres %>%
#   add_slide(master = "Black") %>%
#   ph_with(value = "2019 K-Means Clustering Plot", location = ph_location_type(type = "title")) %>%
#   ph_with(external_img("km2019plot.jpg"), location = ph_location_type(type = "body"), use_loc_size = TRUE)
# 
# UofArkPresNew <- UofArkPres %>%
#   add_slide(master = "Black") %>% 
#   ph_with(value = "2019 K-Means Clustering Data", location = ph_location_type(type = "title")) %>% 
#   ph_with(TeamData2019DTFT, location = ph_location(left = 0.8, top = 2.5))

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

TeamData2019DTFT

```

## K Means clustering 2018

```{r KMeans Cluster 2018, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

TeamDataKM2018 <- scale(TeamDataKM2018)

fit <- kmeans(TeamDataKM2018, 2)

TeamDataKM2018 <- data.frame(TeamDataKM2018, fit$cluster)

TeamData2018 <- data.frame(TeamData, fit$cluster)

rownames(TeamData2018) <- paste(TeamData2018$Sport,TeamData2018$Team,sep="")

clusplot(TeamData2018, fit$cluster, color=TRUE, shade=FALSE, 
  	labels=2, lines=0, cex.txt = .5,col.clus = c("red", "green"),col.txt = c("black", "black"))

TeamData2018DT <- TeamData2018 %>% 
  group_by(TeamData2018$fit.cluster) %>% 
  summarise(count=n(),MeanX2014HF = digits(mean(X2014HF),1),MeanX2015HF = digits(mean(X2015HF),1),MeanX2016HF = digits(mean(X2016HF),1),MeanX2017HF = digits(mean(X2017HF),1),MeanX2018HF = digits(mean(X2018HF),1),MeanPopulation = digits(mean(PopulationMil),1),MeanSalary = digits(mean(X2018Salary),1),MeanESPNRating = digits(mean(ESPNRating),1)) %>%   flextable() %>%
  bold(part = "header") %>% 
  fontsize(size = 12, part = "all") %>% 
  rotate(rotation = "tbrl", align = "center", part = "header") %>% 
  align(align = "right", part = "header") %>% 
  valign(valign = "center", part = "header") %>% 
  align(align = "center", part = "body") %>% 
  set_table_properties(width = .2, layout = "autofit") %>% 
#  autofit() %>% 
  height(height = 1.5, part = "header") %>% 
  bg(bg = "white",part = "all") %>% 
  bg(j = 10, bg = "yellow", part = "all") %>% 
  bg(j = 8:9, bg = "green", part = "all")

TeamData2018DT


```

## K-Means Clustering 2017

```{r KMeans Cluster 2017, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

TeamDataKM2017 <- scale(TeamDataKM2017)

fit <- kmeans(TeamDataKM2017, 2)

TeamDataKM2017 <- data.frame(TeamDataKM2017, fit$cluster)

TeamData2017 <- data.frame(TeamData, fit$cluster)

rownames(TeamData2017) <- paste(TeamData2017$Sport,TeamData2017$Team,sep="")

clusplot(TeamData2017, fit$cluster, color=TRUE, shade=FALSE, 
  	labels=2, lines=0, cex.txt = .5,col.clus = c("red", "green"),col.txt = c("black", "black"))

TeamData2017DT <- TeamData2017 %>% 
  group_by(TeamData2017$fit.cluster) %>% 
  summarise(count=n(),MeanX2014HF = digits(mean(X2014HF),1),MeanX2015HF = digits(mean(X2015HF),1),MeanX2016HF = digits(mean(X2016HF),1),MeanX2017HF = digits(mean(X2017HF),1),MeanPopulation = digits(mean(PopulationMil),1),MeanSalary = digits(mean(X2017Salary),1),MeanESPNRating = digits(mean(ESPNRating),1)) %>% 
  flextable() %>%
  bold(part = "header") %>% 
  fontsize(size = 12, part = "all") %>% 
  rotate(rotation = "tbrl", align = "center", part = "header") %>% 
  align(align = "right", part = "header") %>% 
  valign(valign = "center", part = "header") %>% 
  align(align = "center", part = "body") %>% 
  set_table_properties(width = .2, layout = "autofit") %>% 
#  autofit() %>% 
  height(height = 1.5, part = "header") %>% 
  bg(bg = "white",part = "all") %>% 
  bg(j = 9, bg = "yellow", part = "all") %>% 
  bg(j = 7:8, bg = "green", part = "all")

TeamData2017DT

```

## K Means clustering 2016

```{r KMeans Cluster 2016, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

TeamDataKM2016 <- scale(TeamDataKM2016)

fit <- kmeans(TeamDataKM2016, 2)

TeamDataKM2016 <- data.frame(TeamDataKM2016, fit$cluster)

TeamData2016 <- data.frame(TeamData, fit$cluster)

rownames(TeamData2016) <- paste(TeamData2016$Sport,TeamData2016$Team,sep="")

clusplot(TeamData2016, fit$cluster, color=TRUE, shade=FALSE, 
  	labels=2, lines=0, cex.txt = .5,col.clus = c("red", "green"),col.txt = c("black", "black"))

TeamData2016DT <- TeamData2016 %>% 
  group_by(TeamData2016$fit.cluster) %>% 
  summarise(count=n(),MeanX2014HF = digits(mean(X2014HF),1),MeanX2015HF = digits(mean(X2015HF),1),MeanX2016HF = digits(mean(X2016HF),1),MeanPopulation = digits(mean(PopulationMil),1),MeanSalary = digits(mean(X2016Salary),1),MeanESPNRating = digits(mean(ESPNRating),1)) %>% 
  flextable() %>%
  bold(part = "header") %>% 
  fontsize(size = 12, part = "all") %>% 
  rotate(rotation = "tbrl", align = "center", part = "header") %>% 
  align(align = "right", part = "header") %>% 
  valign(valign = "center", part = "header") %>% 
  align(align = "center", part = "body") %>% 
  set_table_properties(width = .2, layout = "autofit") %>% 
#  autofit() %>% 
  height(height = 1.5, part = "header") %>% 
  bg(bg = "white",part = "all") %>% 
  bg(j = 8, bg = "yellow", part = "all") %>% 
  bg(j = 6:7, bg = "green", part = "all")

TeamData2016DT

```

## K Means clustering 2015

```{r KMeans Cluster 2015, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

TeamDataKM2015 <- scale(TeamDataKM2015)

fit <- kmeans(TeamDataKM2015, 2)

TeamDataKM2015 <- data.frame(TeamDataKM2015, fit$cluster)

TeamData2015 <- data.frame(TeamData, fit$cluster)

rownames(TeamData2015) <- paste(TeamData2015$Sport,TeamData2015$Team,sep="")

clusplot(TeamData2015, fit$cluster, color=TRUE, shade=FALSE, 
  	labels=2, lines=0, cex.txt = .5,col.clus = c("red", "green"),col.txt = c("black", "black"))

TeamData2015DT <- TeamData2015 %>% 
  group_by(TeamData2015$fit.cluster) %>% 
  summarise(count=n(),MeanX2014HF = digits(mean(X2014HF),1),MeanX2015HF = digits(mean(X2015HF),1),MeanPopulation = digits(mean(PopulationMil),1),MeanSalary = digits(mean(X2015Salary),1),MeanESPNRating = digits(mean(ESPNRating),1)) %>%   flextable() %>%
  bold(part = "header") %>% 
  fontsize(size = 12, part = "all") %>% 
  rotate(rotation = "tbrl", align = "center", part = "header") %>% 
  align(align = "right", part = "header") %>% 
  valign(valign = "center", part = "header") %>% 
  align(align = "center", part = "body") %>% 
  set_table_properties(width = .2, layout = "autofit") %>% 
#  autofit() %>% 
  height(height = 2, part = "header") %>% 
  bg(bg = "white",part = "all") %>% 
  bg(j = 7, bg = "yellow", part = "all") %>% 
  bg(j = 5:6, bg = "green", part = "all")

TeamData2015DT

```

```{r K-Means over time, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

ESPNAnalytics <- data.frame("Year" = c(2015,2016,2017,2018,2019,2020),"High" = c(3.5,3.6,3.8,4.0,4.0,4.0),"Low" = c(3.1,3.0,2.9,2.9,2.9,2.9))

ESPNAnalyticsTime <- ESPNAnalytics %>% ggplot() +
  geom_line(aes(x=Year, y=High), color="green", size=1.5) +
  geom_line(aes(x=Year, y=Low), color="red", size=1.5) +
  scale_x_continuous(breaks = seq(2015,2020, by=1)) +
  scale_y_continuous(limits = c(0,6)) +
  ylab("Mean K-Means Group ESPNAnalytics Score") +
  theme(legend.position="none", text = element_text(size = 50)) +
  ggtitle("2015-2020 Gap Between High and Low ESPN Analytics Groups") +
  theme_dark() +
  geom_label(aes(x = Year,y = High, label= paste(digits(High,1))), bg = "green") +
  geom_label(aes(x = Year,y = Low, label= paste(digits(Low,1))), bg = "red", color = "white")

UofArkPresNew <- UofArkPres %>%
  add_slide(master = "Black") %>% 
#  ph_with(value = "2014-2020 Group 1 vs Group 2", location = ph_location_type(type = "title")) %>% 
  ph_with(ESPNAnalyticsTime, location = ph_location_fullsize(), pointsize = 50)

# UofArkPresNew <- UofArkPresNew %>% move_slide(index = 15, to = 9)

ESPNAnalyticsTime


```

```{r Reorder Slides and Produce New Slide Deck, eval = FALSE}

print(UofArkPresNew, target = "Before Slide Moves - U of Ark.pptx")

# for(i in 1:19){
#   UofArkPresNew <- move_slide(UofArkPresNew, index = i+18, to = i+10)
# }
# 
# print(UofArkPresNew, target = "220210 2021 - Is Analytics in Professional Sports Making a Difference - U of Ark.pptx")

```
